sdmTMB: Fast, flexible, and user-friendly spatial and spatiotemporal generalized linear mixed-effects models

### 2021-12-11 doc/paper.Rmd

We can then fit a model with spatial random fields via the function `sdmTMB()`. 
We use a penalized smoother for depth as a main effect via `s()` from the mgcv package.
We specify the family as Tweedie to account for positive continuous density values that also contain zeros.

```{r pcod-eg1-fit, warning=FALSE, message=FALSE, echo=TRUE, eval=FALSE}
fit <- sdmTMB(
  density ~ s(depth),
  data = pcod,
  family = tweedie(link = "log"),
  mesh = mesh,
  spatial = "on"
)
```

@@

```{}
#> Spatiotemporal model fit by ML ['sdmTMB']
#> Formula: density ~ s(depth)
#> Mesh: mesh
#> Data: pcod
#> Family: tweedie(link = 'log')
#>             coef.est coef.se
#> (Intercept)     2.37    0.21
#>
#> Dispersion parameter: 12.69
#> Tweedie p: 1.58
#> Matern range: 16.39
#> Spatial SD: 1.86
#> ML criterion at convergence: 6402.136
```

The output indicates our model was fit by maximum (marginal) likelihood (`'ML'`).
We also see the formula, mesh, fitted data, and family.
Next we see any estimated fixed effects, the Tweedie dispersion and power parameters, the Matérn range distance, the marginal spatial field standard deviation, and the negative log likelihood at convergence.
Internally, `s(depth)` creates random effect spline weights and so is not shown here.

@@

```{r fit-tv, eval=FALSE, echo=TRUE}
fit_spatiotemporal <- sdmTMB(
  density ~ s(depth), family = tweedie(link = "log"), data = pcod, mesh = mesh,
  time = "year", spatial = "off", spatiotemporal = "ar1"
)
```

>>>>

We can then fit a model with spatial random fields via the function `sdmTMB()`. 
We use a penalized smoother for depth as a main effect via `s()` from the mgcv package and choose a basis dimension of 5.
We specify the family as Tweedie to account for positive continuous density values that also contain zeros.

```{r pcod-eg1-fit, warning=FALSE, message=FALSE, echo=TRUE, eval=FALSE}
fit <- sdmTMB(
  density ~ s(depth, k = 5),
  data = pcod,
  family = tweedie(link = "log"),
  mesh = mesh,
  spatial = "on"
)
```

@@

```{}
#> Spatiotemporal model fit by ML ['sdmTMB']
#> Formula: density ~ s(depth, k = 5)
#> Mesh: mesh
#> Data: pcod
#> Family: tweedie(link = 'log')
#>             coef.est coef.se
#> (Intercept)     2.34    0.23
#> sdepth        -26.96   21.52
#>
#> Smooth terms:
#>            Std. Dev.
#> sds(depth)     34.03
#>
#> Dispersion parameter: 12.70
#> Tweedie p: 1.58
#> Matern range: 17.47
#> Spatial SD: 1.86
#> ML criterion at convergence: 6403.637
```

The output indicates our model was fit by maximum (marginal) likelihood (`ML`).
We also see the formula, mesh, fitted data, and family.
Next we see any estimated main effects including the linear component of the smoother (`sdepth`), the standard deviation on the smoother weights (`sds(depth)`), the Tweedie dispersion and power parameters, the Matérn range distance, the marginal spatial field standard deviation, and the negative log likelihood at convergence.

@@

```{r fit-tv, eval=FALSE, echo=TRUE}
fit_spatiotemporal <- sdmTMB(
  density ~ s(depth, k = 5), family = tweedie(link = "log"), data = pcod, mesh = mesh,
  time = "year", spatial = "off", spatiotemporal = "ar1"
)
```
